Peer-graded Assignment: Course Project 1
========================================================

Load the data

```{r}
activityData <- read.csv("activity.csv")
head(activityData)
```
# What is mean total number of steps taken per day?

calculate total steps per day
```{r}
StepsPerDay<-tapply(activityData$steps,activityData$date,sum) 

#convert stepsPerDay to a dataframe
StepsPerDay<-data.frame(StepsPerDay) 
names(StepsPerDay)<-c("steps")
```
Make a histogram of the total number of steps taken each day
```{r}
stepsFreq<-hist(StepsPerDay$steps,main = "Histogram of total number of steps per day", xlab = "Steps per day")
```

Calculate mean and median of the total number of steps taken per day
```{r}
summary(StepsPerDay)
```
Median :10765, Mean   :10766

# What is the average daily activity pattern?
```{r}
interval_steps <- aggregate(steps ~ interval, activityData, mean)

plot(interval_steps$interval, interval_steps$steps, type='l', col=1, 
     main="Average number of steps averaged across all days", xlab="Interval", 
     ylab="Average number of steps")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps
```{r}
MaxNumOfSteps <- which.max(interval_steps$steps)

# get the interval with maximum average number of steps in an interval
interval_steps [MaxNumOfSteps, ]
```
# Imputing missing values
```{r}
#Calculate the total number of missing values in the dataset
sum(!complete.cases(activityData))
```
replace na values with the mean for that 5-minute interval
```{r}
for (i in 1:nrow(activityData)) {
        if (is.na(activityData$steps[i])) {
                Corresponding_interval<-activityData$interval[i]
                activityData$steps[i]<-interval_steps$steps[interval_steps$interval==Corresponding_interval]
        }   
}
#Create a new dataset that is equal to the original dataset but with the missing data filled in.
new_activityData<-activityData

#Make a histogram of the total number of steps taken each day
new_StepsPerDay<-tapply(new_activityData$steps,new_activityData$date,sum)  
new_StepsPerDay<-data.frame(new_StepsPerDay)
names(new_StepsPerDay)<-c("steps")
new_stepsFreq<-hist(new_StepsPerDay$steps,main = "Histogram of total number of steps per day", xlab = "Steps per day")

# get the mean and meadian
summary(new_StepsPerDay)
```
there is a slight difference in meadian value.
# Are there differences in activity patterns between weekdays and weekends?
```{r}
# Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
day <- weekdays(as.Date(new_activityData$date))
day_type <- vector()
for (i in 1:nrow(new_activityData)) {
        if (day[i] == "Saturday") {
                day_type[i] <- "Weekend"
        } else if (day[i] == "Sunday") {
                day_type[i] <- "Weekend"
        } else {
                day_type[i] <- "normalday"
        }
}
new_activityData$day_type <- day_type

# convert day_type from character to factor
new_activityData$day_type <- as.factor(new_activityData$day_type)

#Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 
new_interval_steps <- aggregate(steps ~ interval+day_type, new_activityData, mean)
library(lattice)

# create the panel plot
xyplot(steps ~ interval | day_type, new_interval_steps, type = "l", layout = c(1, 2), 
       xlab = "interval", ylab = "Number of steps")
```

